% ************************** Thesis Acknowledgements **************************

\begin{acknowledgements}      

First and foremost, I would like to thank my supervisor Dr. Igor
Mekhov who has been an excellent mentor throughout my time at
Oxford. It is primarily thanks to his brilliant insights and
professionalism that I was able to reach my full potential during my
doctoral studies. The work contained in this thesis would also not be
possible without the help of the other members of the group, Gabriel
Mazzucchi and Dr. Santiago Caballero-Benitez. Without our frequent
casual discussions in the Old Library office I would have still been
stuck on the third chapter. I would also like to acknowledge all
members of Prof. Dieter Jaksch's and Prof. Christopher Foot's groups
for various helpful discussions. I must also offer a special mention
for Edward Owen who provided much needed reality checks on some of my
wishful theoretical thinking. I would also like to express my
gratitude to EPSRC, St. Catherine's College, the ALP sub-department,
and the Institute of Physics for providing me with the financial means
to live and study in Oxford as well as attend several conferences in
the UK and abroad.

On a personal note, I would like to thank my parents who provided me
with all the skills necessary work towards any goals I set
myself. Needless to say, without them I would have never been able to
be where I am right now. I would like to thank all the new and old
friends that have kept me company for the last four years. The time
spent together provided a welcome respite from the sweat and toil of
my DPhil.

\vspace{2em}
\raggedleft{Oxford, September 2016}

\end{acknowledgements}
